Peer Assessment 1
========================================================

## Loading and preprocessing the data

```{r}
library(dplyr)
library(lattice)
options(scipen=10)
data <- read.csv(paste0(getwd(),"/activity.csv"))
by_day <- group_by(data, date)
total_steps <- summarize(by_day, total = sum(steps))
total_steps$date <- as.Date(total_steps$date, "%Y-%m-%d")

```

## What is mean total number of steps taken per day?

```{r fig.width=7, fig.height=6}

hist(total_steps$total, breaks=20,xlab="Total number of steps", main="Total number of steps per day")
```

```{r}
meansteps <- mean(total_steps$total, na.rm=T)
mediansteps <- median(total_steps$total, na.rm=T)
```

On average, a total number of `r round(meansteps,2)` has been recorded. The median of the total number of steps equals `r mediansteps` .

## What is the average daily activity pattern?

```{r fig.width=7, fig.height=6}
by_5min <- group_by(data, interval)
mean_steps <- summarise(by_5min, mean=mean(steps,na.rm=T))

plot(mean_steps$interval, mean_steps$mean, type="l", main="Average      number of steps per interval", xlab="Time interval", ylab="Average number of steps")
max_interval <- mean_steps[which.max(mean_steps$mean),]

```

The maximum average number of steps, `r round(max_interval[2],2)`, occures at the `r round(max_interval[1])`-interval in the data.


## Imputing missing values

```{r}
colSums(is.na(data))
```

A simple way of imputing missing values is to use the average number of steps of the corresponding interval.


```{r}
data.na.rm <- data
for(i in 1:dim(data)[1]) {
if(is.na(data[i,1])) {data.na.rm[i,1] <- mean_steps[which(mean_steps$interval==data[i,3]),2]}
}
```

```{r fig.width=7, fig.height=6}
by_dayr <- group_by(data.na.rm, date)
total_stepsr <- summarize(by_dayr, total = sum(steps, na.rm=T))
hist(total_stepsr$total, breaks=20,xlab="Total number of steps", main="Total number of steps per day with NA correction")
```

```{r}
options(scipen=3)
meanstepsr <- mean(total_stepsr$total)
medianstepsr <- median(total_stepsr$total)
```

With correction for the missing values, the mean is `r round(meanstepsr,2)` and the median equals `r round(medianstepsr,2)`.  
Hence, imputing missing values does not change the mean but the median increases about `r round(medianstepsr-mediansteps,2)` and is now equal to the mean. 


## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekd <- mutate(data.na.rm, weekend = factor(weekdays(as.Date(date,"%Y-%m-%d"))=="Samstag" | weekdays(as.Date(date,"%Y-%m-%d"))=="Sonntag", labels=c("weekdays","weekend")))

by_5min_weekd <- group_by(weekd, interval, weekend)
mean_steps_weekd <- summarise(by_5min_weekd, mean=mean(steps,na.rm=T))


xyplot(mean ~ interval|weekend, mean_steps_weekd, type="l", layout=c(1,2), xlab="Time interval", ylab="Average number of steps per  interval")

```




